---
title: "Data Preprocessing"
author: "Group Rhapsody"
date: "26 Dec, 2020"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: true
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Importing Necessary Libraries 

```{r, results='hide', message=FALSE}
library(tidyverse)
library(rio)
library(knitr)
```

### Reading & Preparing Datasets

In the project, we have four main datasets that we used to make analysis. Below, you can find the reading and preparing stages of these datasets.

#### Dataset 1 - Unemployed Job Searching People by the Channel

The original dataset can be found at [this link](https://data.tuik.gov.tr/Bulten/DownloadIstatistikselTablo?p=YD6JGk3HAkwb1wQadeYsjp8iMZJ1cmLtRnMCRXPG3aOudsWPijD1R0XxxA54Ehu8).

The dataset is grouped by gender and includes job search methods. The data is between January 2014 and August 2020.

+ Since the data in the file does not start with A1 cell, the ranges are given as parameters during the import process.
+ All columns are renamed accordingly during the import process.
+ The original dataset has some empty cells for the `year` column. Therefore, empty cells filled with the correct year value.
+ The data table converted to a tibble.
+ `month` column includes both Turkish and English month names. By using `str_split_fixed` function, only the English name is taken.

##### All Genders Dataset

```{r}
job_search_overall <- import("https://github.com/pjournal/mef04g-rhapsody/blob/gh-pages/Project_Data/search_channel.xls?raw=true", 
                             range = "TÜRKİYE!A6:P91", 
                             col_names = c("year", "month", "total_unemployed", "to_employers", "to_relatives", "to_emp_office", "to_emp_agencies", "to_newspaper", "insert_ad_to_newspaper", "take_interview", "look_place_equip_to_est_bus", "look_credit_license_to_est_bus", "wait_call_from_emp_office", "wait_result_of_app", "wait_result_of_comp_for_public_sec", "others")) %>% 
  fill(year, .direction = "down") %>%
  as_tibble()

job_search_overall$month <- str_split_fixed(job_search_overall$month, " - ", 2)[,2]
```

Final tibble is as follows.

```{r, echo=FALSE}
glimpse(job_search_overall)
```


##### Male Dataset

```{r}
job_search_male <- import("https://github.com/pjournal/mef04g-rhapsody/blob/gh-pages/Project_Data/search_channel.xls?raw=true", 
                          range = "TÜRKİYE!R6:AG91", 
                          col_names = c("year", "month", "total_unemployed", "to_employers", "to_relatives", "to_emp_office", "to_emp_agencies", "to_newspaper", "insert_ad_to_newspaper", "take_interview", "look_place_equip_to_est_bus", "look_credit_license_to_est_bus", "wait_call_from_emp_office", "wait_result_of_app", "wait_result_of_comp_for_public_sec", "others")) %>% 
  fill(year, .direction = "down") %>%
  as_tibble()

job_search_male$month <- str_split_fixed(job_search_male$month, " - ", 2)[,2]
```

Final tibble is as follows.

```{r, echo=FALSE}
glimpse(job_search_male)
```

##### Female Dataset

```{r}
job_search_female <- import("https://github.com/pjournal/mef04g-rhapsody/blob/gh-pages/Project_Data/search_channel.xls?raw=true", 
                            range = "TÜRKİYE!AI6:AX91", 
                            col_names = c("year", "month", "total_unemployed", "to_employers", "to_relatives", "to_emp_office", "to_emp_agencies", "to_newspaper", "insert_ad_to_newspaper", "take_interview", "look_place_equip_to_est_bus", "look_credit_license_to_est_bus", "wait_call_from_emp_office", "wait_result_of_app", "wait_result_of_comp_for_public_sec", "others")) %>% 
  fill(year, .direction = "down") %>%
  as_tibble()

job_search_female$month <- str_split_fixed(job_search_female$month, " - ", 2)[,2]
```

Final tibble is as follows.

```{r, echo=FALSE}
glimpse(job_search_female)
```

#### Dataset 2 - Employed & Unemployed by Educational Level

The original dataset can be found at [this link](https://data.tuik.gov.tr/Bulten/DownloadIstatistikselTablo?p=bWJMnEy0j6UZHm1gD5qqGkGNWSNKDEJ4I57rthEhkNd9qHF6QHUpwQl25NrXRpAg).

This dataset includes the number of employed and unemployed people by their educational levels and gender. The data is between January 2014 and August 2020.

+ Since the data in the file does not start with A1 cell, the ranges are given as parameters during the import process.
+ All columns are renamed accordingly during the import process.
+ There are two empty columns in the file. These empty columns are eliminated by using `select` function.
+ The original dataset has some empty cells for the `year` column. Therefore, empty cells filled with the correct year value.
+ The data table converted to a tibble.
+ `month` column includes both Turkish and English month names. By using `str_split_fixed` function, only the English name is taken.
+ There is a difference in two January values in the `month` column which causes these values to be equal to empty strings. Therefore, these values are being found and updated as January.

##### All Genders Dataset

```{r}
educational_level_overall <- import("https://github.com/pjournal/mef04g-rhapsody/blob/gh-pages/Project_Data/educational_level.xls?raw=true", 
                                    range = "TÜRKİYE!A7:S92", 
                                    col_names = c("year", "month", "lf_illeterate", "lf_less_than_hs", "lf_highschool", "lf_voc_hs", "lf_higher_ed", "empty_col_1", "emp_illeterate", "emp_less_than_hs", "emp_highschool", "emp_voc_hs", "emp_higher_ed", "empty_col_2", "unemp_illeterate", "unemp_less_than_hs", "unemp_highschool", "unemp_voc_hs", "unemp_higher_ed" )) %>% 
  select(-empty_col_1, -empty_col_2) %>% 
  fill(year, .direction = "down") %>%
  as_tibble()

educational_level_overall$month <- str_split_fixed(educational_level_overall$month, " - ", 2)[,2]

educational_level_overall$month[educational_level_overall$month == ""] = "January"
```

Final tibble is as follows.

```{r, echo=FALSE}
glimpse(educational_level_overall)
```



#### Dataset 3 - Unemployment by Occupational Group

The original dataset can be found at [this link](https://data.tuik.gov.tr/Bulten/DownloadIstatistikselTablo?p=VyBwucwx3Wx2uJXGxoM3soWzPxQv5BRhYRELMU0FhptFZ5Ythna2czyzH6s4rDxR).

This dataset includes the number of unemployed people by their occupational group and gender. The data is between January 2014 and August 2020.

+ Since the data in the file does not start with A1 cell, the ranges are given as parameters during the import process.
+ All columns are renamed accordingly during the import process.
+ The original dataset has some empty cells for the `year` column. Therefore, empty cells filled with the correct year value.
+ The data table converted to a tibble.
+ `month` column includes both Turkish and English month names. By using `str_split_fixed` function, only the English name is taken.

##### All Genders Dataset

```{r}
occ_group_overall <- import("https://github.com/pjournal/mef04g-rhapsody/blob/gh-pages/Project_Data/occupational_group.xls?raw=true", 
                            range = "TÜRKİYE!A7:L92", 
                            col_names = c("year", "month", "total_unemployed", "manager", "prof", "tech", "cleric", "service", "agricul", "trade", "operator", "elemantary")) %>%
  fill(year, .direction = "down") %>%
  as_tibble()

occ_group_overall$month <- str_split_fixed(occ_group_overall$month, " - ", 2)[,2]
```

Final tibble is as follows.

```{r, echo=FALSE}
glimpse(occ_group_overall)
```

##### Male Dataset

```{r}
occ_group_male <- import("https://github.com/pjournal/mef04g-rhapsody/blob/gh-pages/Project_Data/occupational_group.xls?raw=true", 
                         range = "TÜRKİYE!N7:Y92", 
                         col_names = c("year", "month", "total_unemployed", "manager", "prof", "tech", "cleric", "service", "agricul", "trade", "operator", "elemantary")) %>%
  fill(year, .direction = "down") %>%
  as_tibble()

occ_group_male$month <- str_split_fixed(occ_group_male$month, " - ", 2)[,2]
```

Final tibble is as follows.

```{r, echo=FALSE}
glimpse(occ_group_male)
```

##### Female Dataset

```{r}
occ_group_female <- import("https://github.com/pjournal/mef04g-rhapsody/blob/gh-pages/Project_Data/occupational_group.xls?raw=true", 
                           range = "TÜRKİYE!AA7:AL92", 
                           col_names = c("year", "month", "total_unemployed", "manager", "prof", "tech", "cleric", "service", "agricul", "trade", "operator", "elemantary")) %>%
  fill(year, .direction = "down") %>%
  as_tibble()

occ_group_female$month <- str_split_fixed(occ_group_female$month, " - ", 2)[,2]
```

Final tibble is as follows.

```{r, echo=FALSE}
glimpse(occ_group_female)
```

#### Dataset 4 - Labour Force Status of Higher Education Graduates by Major

The original dataset can be found at [this link](https://data.tuik.gov.tr/Bulten/DownloadIstatistikselTablo?p=vLICiE6Ol0yQYtmtU/kJ2JEnAr8Q14emL5iDhwHMpzChnPnzsVhz1tXV3REhaV9R).

It consists of annual numbers of employed and unemployed people by field of education. The dataset consists of the statistics from 2014 to 2019

+ Since the data in the file does not start with A1 cell, the ranges are given as parameters during the import process.
+ All columns are renamed accordingly during the import process.
+ The original dataset has some empty cells for the `year` column. Therefore, empty cells filled with the correct year value.
+ The data table converted to a tibble.

##### Dataset

```{r}
last_graduated_major <- import("https://github.com/pjournal/mef04g-rhapsody/blob/gh-pages/Project_Data/major_field.xls?raw=true", 
                               range = "TURKIYE!A6:Y41", 
                               col_names = c("year", "statistics", "higher_ed_grad", "education", "arts", "humanities", "languages", "social_sci", "journalism", "business", "law", "biology_env_related_sci", "physical_sci", "math_stat", "info_communication_tech", "engineering", "manufacturing_processing", "architecture_construction", "agriculture_forestry_fishery", "veterinary", "health", "welfare", "personal_services", "occupational_health_transport_services", "security_services")) %>%
  fill(year, .direction = "down") %>%
  as_tibble()
```

Final tibble is as follows.

```{r, echo=FALSE}
glimpse(last_graduated_major)
```

### Creating the .Rdata File

After preparing all the datasets, we can create a .Rdata file. In order to do this, all tibbles are saved into a single file named `project_all_data.RData`. In further analysis, loading this .Rdata file will be sufficient to reach all the necessary data.

```{r, eval=FALSE}
save(job_search_overall, job_search_male, job_search_female, educational_level_overall,
     occ_group_overall, occ_group_male, occ_group_female, last_graduated_major, 
     file = "project_all_data.RData") 
```

### References

+ [Reading .xls Files From URL](https://github.com/tidyverse/readxl/issues/183)
+ [Creating .Rdata Files](https://bookdown.org/ndphillips/YaRrr/rdata-files.html)
+ [Filling Empty Values with Previous Values in a Column](https://stackoverflow.com/questions/40040834/replace-na-with-previous-or-next-value-by-group-using-dplyr/40041172)

Back to [our progress journal.](https://pjournal.github.io/mef04g-rhapsody/)