---
title: "Rhapsody Group Project"
subtitle: "Unemployment & Employment in 2014-2020"
author: "Cem Yüzbaşıoğlu<br/>Emre Demirhan<br/>Emre Kemal Yurderi<br/>Serhan Bayram<br/>Umut Turhan"
institute: "MEF University"
date: "30 Dec, 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(lubridate)
library(reshape2)
library(knitr)
library(gridExtra)
```

```{r, echo=FALSE}
load(url("https://github.com/pjournal/mef04g-rhapsody/blob/gh-pages/Project_Data/project_all_data.RData?raw=true"))
```

# Introduction

We used four different datasets which are obtained from TÜIK.

+ Unemployed People by the Job Searching Channel<sup>1</sup>
+ Employed & Unemployed People by Educational Level<sup>2</sup>
+ Unemployment by Occupational Group<sup>3</sup>
+ Unemployment of Higher Education Graduates by Major<sup>4</sup>

We analysed each dataset and combined different outcomes.

.footnote[
[1] [Dataset 1 - Link](https://data.tuik.gov.tr/Bulten/DownloadIstatistikselTablo?p=YD6JGk3HAkwb1wQadeYsjp8iMZJ1cmLtRnMCRXPG3aOudsWPijD1R0XxxA54Ehu8)

[2] [Dataset 2 - Link](https://data.tuik.gov.tr/Bulten/DownloadIstatistikselTablo?p=bWJMnEy0j6UZHm1gD5qqGkGNWSNKDEJ4I57rthEhkNd9qHF6QHUpwQl25NrXRpAg)

[3] [Dataset 3 - Link](https://data.tuik.gov.tr/Bulten/DownloadIstatistikselTablo?p=VyBwucwx3Wx2uJXGxoM3soWzPxQv5BRhYRELMU0FhptFZ5Ythna2czyzH6s4rDxR)

[4] [Dataset 4 - Link](https://data.tuik.gov.tr/Bulten/DownloadIstatistikselTablo?p=vLICiE6Ol0yQYtmtU/kJ2JEnAr8Q14emL5iDhwHMpzChnPnzsVhz1tXV3REhaV9R)
]
---

## Unemployment in 2014-2020

```{r, echo=FALSE, fig.width=10, fig.height=3.5}
occ_group_overall_m <- occ_group_overall %>%
  filter(month != "Annual") %>%
  mutate(year = factor(year),
         month = factor(month, levels = c("January","February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")))

occ_group_overall_m <- occ_group_overall_m %>%
  mutate(date = factor(paste(year, month), levels = paste(year, month)))

unemployment_all <- ggplot(occ_group_overall_m, aes(date, total_unemployed)) +
  geom_line(aes(group = 1)) +
  labs(x = "",
       y = "Unemployment (in thousands)") +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_text(angle = 90, size = 6,  hjust = 1)) +
  geom_segment(aes(x = 31, y = 3000, xend = 36, yend = 3600)) +
  geom_segment(aes(x = 31, y = 3000, xend = 36, yend = 3600),
               color="red",
               arrow = arrow(length = unit(0.3, "cm"))) +
  geom_segment(aes(x = 56, y = 3200, xend = 62, yend = 4400)) +
  geom_segment(aes(x = 56, y = 3200, xend = 62, yend = 4400),
               color="red",
               arrow = arrow(length = unit(0.3, "cm")))

unemployment_all
```

--

```{r, echo=FALSE, fig.width=10, fig.height=3.5}
monthly_jobsearch_male <- filter(job_search_male, !grepl("Annual", month))
monthly_jobsearch_female <- filter(job_search_female, !grepl("Annual", month))

unemp_gr_male<-monthly_jobsearch_male%>%
  transmute(month_year = paste(year, month, sep = " "), total_unemployed)

unemp_gr_female<-monthly_jobsearch_female%>%
  transmute(month_year = paste(year, month, sep = " "), total_unemployed)


unemp_merged <-merge(unemp_gr_female,unemp_gr_male, by="month_year", all=TRUE, sort = FALSE)

unemp_merged <- unemp_merged %>%
  rename(female=total_unemployed.x, male=total_unemployed.y)

unemp_merged$month_year <- as.character(unemp_merged$month_year)
unemp_merged$month_year <- factor(unemp_merged$month_year, level=unemp_merged$month_year)

unemployment_gender <- ggplot(unemp_merged, aes(x=factor(month_year))) +
  geom_line(aes(y=female, color="Female", group=1))+
  geom_line(aes(y=male, color="Male", group=1))+
  theme(axis.text.x = element_text(angle=90, size=5, hjust = 1),
        plot.title = element_text(hjust = 0.5))+
  labs(x="",
       y="Unemployment (in thousands)",
       color="Gender")

unemployment_gender
```

---

## Employed & Unemployed by Educational Level

```{r, include=FALSE}
yearly_educational <- filter(educational_level_overall, grepl("Annual", month))
total_labour_yearly_educational <- yearly_educational%>%
  transmute(year=as.character(year),illeterate=lf_illeterate, less_than_hs=lf_less_than_hs, highschool=lf_highschool, voc_hs=lf_voc_hs, higher_ed=lf_higher_ed)

total_labour_yearly_educational_melt <- total_labour_yearly_educational %>%
  gather(key="type", value="value", -year) %>%
  group_by(year) %>%
  mutate(Percentage= round(value/sum(value) * 100, 1))


emp_yearly_educational <- yearly_educational %>%
  transmute(year=as.character(year), illeterate=emp_illeterate, less_than_hs=emp_less_than_hs, highschool=emp_highschool, voc_hs=emp_voc_hs, higher_ed=emp_higher_ed)

emp_yearly_educational_melt<-emp_yearly_educational %>%
  gather(key="Education_Level", value="Person", -year ) %>%
  group_by(year) %>%
  mutate(Percentage= round(Person/sum(Person) * 100, 1))

emp_yearly_educational_melt$Education_Level <- as.factor(emp_yearly_educational_melt$Education_Level)
levels(emp_yearly_educational_melt$Education_Level) <- c("Higher Education", "Highschool", "Illeterate", "Less Than Highschool", "Vocational Highschool")


unemp_yearly_educational <- yearly_educational %>%
  transmute(year=as.character(year), illeterate=unemp_illeterate, less_than_hs=unemp_less_than_hs, highschool=unemp_highschool, voc_hs=unemp_voc_hs, higher_ed=unemp_higher_ed)

unemp_yearly_educational_melt<-unemp_yearly_educational %>%
  gather(key="Education_Level", value="Person", -year ) %>%
  group_by(year) %>%
  mutate(Percentage= round(Person/sum(Person) * 100, 1))

unemp_yearly_educational_melt$Education_Level <- as.factor(unemp_yearly_educational_melt$Education_Level)
levels(unemp_yearly_educational_melt$Education_Level) <- c("Higher Education", "Highschool", "Illeterate", "Less Than Highschool", "Vocational Highschool")
```

```{r, echo=FALSE, fig.width=10, fig.height=3.5}
plot_emp_df_higher <- emp_yearly_educational_melt %>% filter(Education_Level == "Higher Education")

plot_emp_higher <- ggplot(plot_emp_df_higher, aes(x = year, y = Percentage)) +
  geom_bar(stat = "identity", aes(fill = as.factor(Education_Level))) +
  coord_cartesian(ylim=c(0,70))+
  geom_text(aes(label = paste0(Percentage, "%"), vjust = -1), size = 4) +
  labs(title="Employment Rates of Higher Education", x="", y="% of People") +
  theme(legend.position = "none",  plot.title = element_text(hjust = 0.5))


plot_unemp_df_higher <- unemp_yearly_educational_melt %>% filter(Education_Level == "Higher Education")

plot_unemp_higher <- ggplot(plot_unemp_df_higher, aes(x = year, y = Percentage)) +
  geom_bar(stat = "identity", aes(fill = as.factor(Education_Level))) +
  coord_cartesian(ylim=c(0,70))+
  geom_text(aes(label = paste0(Percentage, "%"), vjust = -1), size = 4) +
  facet_wrap(~ Education_Level)+
  labs(title="Unemployment Rates of Higher Education", x="", y="") +
  theme(legend.position = "none",  plot.title = element_text(hjust = 0.5))

grid.arrange(plot_emp_higher, plot_unemp_higher, ncol=2)
```

--

```{r, echo=FALSE, fig.width=10, fig.height=3.5}
plot_emp_df_less <- emp_yearly_educational_melt %>% filter(Education_Level == "Less Than Highschool")

plot_emp_less <- ggplot(plot_emp_df_less, aes(x = year, y = Percentage)) +
  geom_bar(stat = "identity") +
  coord_cartesian(ylim=c(0,70))+
  geom_text(aes(label = paste0(Percentage, "%"), vjust = -1), size = 4) +
  labs(title="Employment Rates of Less Than Highschool", x="", y="% of People") +
  theme(legend.position = "none",  plot.title = element_text(hjust = 0.5))


plot_unemp_df_less <- unemp_yearly_educational_melt %>% filter(Education_Level == "Less Than Highschool")

plot_unemp_less <- ggplot(plot_unemp_df_less, aes(x = year, y = Percentage)) +
  geom_bar(stat = "identity") +
  coord_cartesian(ylim=c(0,70))+
  geom_text(aes(label = paste0(Percentage, "%"), vjust = -1), size = 4) +
  facet_wrap(~ Education_Level)+
  labs(title="Unemployment Rates of Less Than Highschool", x="", y="") +
  theme(legend.position = "none",  plot.title = element_text(hjust = 0.5))


grid.arrange(plot_emp_less, plot_unemp_less, ncol=2)
```

---

## Unemployment of Higher Education Graduates

```{r, echo=FALSE, fig.width=10}
deneme_major <- last_graduated_major %>%
  gather(key = "Major", value = "value", -year, -statistics)

deneme_major_issizoran <- filter(deneme_major, grepl("İşsizlik oranı", statistics))

major_issizoran_wide <- spread(deneme_major_issizoran, year, value) %>%
  subset(select = -c(statistics))

major_issizoran_wide$Trend <- round(((major_issizoran_wide$`2019`/major_issizoran_wide$`2014`)-1)*100,1)

major_issizoran_wide$Major <- as.factor(major_issizoran_wide$Major)
levels(major_issizoran_wide$Major) <- c("Agriculture, forestry and fishery","Architecture and construction", "Arts", "Biology & Env. Sci.", "Business & Admin.", "Education", "Engineering and engineering trades", "Health", "Higher education graduate", "Humanities", "Info. & Comm. Tech.", "Journalism & Info.", "Languages", "Law", "Manufacturing and processing", "Mathematics and statistics", "Occupational health and transport services", "Personal Services", "Physical science", "Security Services", "Social and behavioural sciences", "Veterinary", "Welfare (Social services)")

major_top5 <- major_issizoran_wide %>%
  top_n(5, Trend)

major_last5 <- major_issizoran_wide %>%
  arrange(Trend) %>%
  head(5)

kable(mutate(major_issizoran_wide, across(where(is.numeric), round, 1)) %>% 
        filter(Major %in% c("Health","Journalism & Info.","Personal Services","Biology & Env. Sci.","Info. & Comm. Tech.","Architecture and construction","Engineering and engineering trades","Business & Admin.")))

major_top_plot <- ggplot(major_top5, aes(factor(Major, levels = Major), Trend, fill = Major)) +
  geom_col(position = "dodge") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))  +
  guides(fill=guide_legend(title="Major")) +
  labs(x = "",
       y = "Percentage of Change from 2014 to 2019",
       title = "Top 5 Major by Unemployment Rate Trend")

major_last_plot <- ggplot(major_last5, aes(factor(Major, levels = Major), Trend, fill = Major)) +
  geom_col(position = "dodge") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))  +
  guides(fill=guide_legend(title="Major")) +
  labs(x = "",
       y = "Percentage of Change from 2014 to 2019",
       title = "Lowest 5 Major by Unemployment Rate Trend")
```

---

## Unemployment of Higher Education Graduates

```{r, echo=FALSE, fig.width=10, fig.height=4}
major_last_plot

kable(mutate(major_issizoran_wide, across(where(is.numeric), round, 1)) %>% 
        filter(Major == "Journalism & Info."))
```

---

## Unemployment of Higher Education Graduates

```{r, echo=FALSE, fig.width=10, fig.height=4}
major_top_plot

kable(mutate(major_issizoran_wide, across(where(is.numeric), round, 1)) %>% 
        filter(Major == "Health"))
```

---

## Unemployed Job Searching People by the Channel

```{r, echo=FALSE, fig.width=10}
job_search_overall_c <- job_search_overall %>%
  filter(month == "Annual") %>%
  mutate(year = factor(year), 
         establish_own_bus = look_place_equip_to_est_bus + look_credit_license_to_est_bus,
         newspaper = to_newspaper + insert_ad_to_newspaper,
         iskur = to_emp_office + wait_call_from_emp_office,
         waiting_result = wait_result_of_app + wait_result_of_comp_for_public_sec) %>%
  select(year, establish_own_bus, newspaper, iskur, to_emp_agencies, to_employers, to_relatives)

Year <- rep(unique(job_search_overall_c$year), times = (ncol(job_search_overall_c)-1))

Channel <- rep(colnames(job_search_overall_c[-c(1,1)]),
               each = length(unique(Year)))

Count <- array(unlist(job_search_overall_c[1:6,2:7]))

js_df <- data.frame(Year, Channel, Count)

js_df_rate <- js_df %>%
  group_by(Year) %>%
  mutate(Percentage = round(Count/sum(Count) * 100, 1))

js_df_rate$Channel <- as.factor(js_df_rate$Channel)

levels(js_df_rate$Channel) <- c("Establish Own Business", "Turkish Employment Office", "Newspaper",
                                "Employment Agencies", "Direct Application to Employers", "Friends & Relatives")

ggplot(js_df_rate, aes(x = Year, y = Percentage)) +
  geom_bar(stat = "identity", aes(fill = as.factor(Channel))) +
  geom_text(aes(label = paste0(Percentage, "%"), vjust = -1)) +
  facet_wrap(~Channel) +
  ggtitle("Rates of Job Search Channel Among All Applications") + 
  theme(legend.position = "none",  plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(0,40))
```

---

## Conclusion

+ Number of unemployed in January 2014 is nearly 2.8 million, in August 2020 is nearly 4.2 million. 
An increase of **50%**.

+ **Higher Education** and **Less Than Highschool** levels have the opposite trends both in employment and unemployment.

+ Some safe-considered majors have increasing unemployment rates such as **Health**, **Architecture** and **Engineering**.

+ Increase in the number of job applications through Turkish Employment Office, might be related with the decline in job opportunities.

---

class: center, middle

# Thank You For Listening!

[Check Out **Rhapsody's Progress Journal**](https://pjournal.github.io/mef04g-rhapsody/).

## Project Documents

[Data Preprocessing](https://pjournal.github.io/mef04g-rhapsody/DataPreprocessing.html)

[Final Report](https://pjournal.github.io/mef04g-rhapsody/FinalReport.html)

[Shiny App]()

